{\large \bfseries
Contribution to the methodological re-design of the assessment of sustainability by life cycle thinking, through the explicit integration of value judgements, the treatment of multi-dimensionality and a free and alternative approach to data.
%Operational Holistic Evaluation fondations \\ Methodological re-design of the sustainability assessment by life cycle thinking and artificialism}
}
~\\

\textbf{Abstract:}\\
Humanity goes towards unprecedented challenges in its history.
To guide decisions on a more \emph{sustainable} way various artefacts were produced.
However, those have failed to respond to the sustainability quest.
From the renewed conceptual design of Life Cycle Analysis (LCA) methodology, trough artificialism, derive the foundations of Operational Holistic Evaluation (OHE).

Explicit value judgement integration, essential to rational decisions, combined with multidimensionality study, first with multifunctionality then in its generalization, shaked the established principles of the current life cycle thinking.

Some bases persist~:
The common base of decision canonical form and its iterative formulation with stages of~: intelligence of the problem (goal), creativity (inventory and search for alternatives), choice (interpretation).

But others are eliminated~:
The techno-eco-sphere model and consequently the notions of elementary or product flows, keeping only the concepts of flow and system.
Impacts methods logic, to keep only the description of technical, social and environmental systems supplying the multiple-criteria decision analysis.
The areas of protection are merged with the needs domain (and related functions).
Last but not the least, the academic logic of production and dissemination of knowledge, whose renovation goes hand in hand with the extinction of the current dominations in LCA and generally Science, in favour of a web, semantic, interactive and free.

This thesis, marking the union between Environmental Assessment and Economy, aims at moving from LCA to OHE.

~\\
\textbf{Keywords:}\\
Sustainability~; Life Cycle Assessment~; Multifunctionality; ~ Multidimensionality~; Value judgments; Multicriteria decision analysis~; Design~; Artificialism
\newpage